---
title: "Ethiopia Stochastic Frontiers Analysis Draft 2"
author: "Tomas Morley, Michiel van Dijk, Assefa Admassie, Roel Jongeneel"
date: "30 March 2017"
output:
  pdf_document: default
  word_document:
    reference_docx: ../Common/word_styles_01.docx
bibliography: ../Common/ETHYG.bib
---

```{r setup, include=FALSE}
# Knitr settings
library(rprojroot)
library(dplyr)
root <- find_root(is_rstudio_project)

library(frontier)
library(knitr)
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"FigTabMap/out-"),
  dev = "CairoPNG",
  dev.args = list(CairoPNG = list(bg = "transparent"))
  )

library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

source(file.path(root, "Code/production_function_tests.R"))

# Source figures, tables and maps
source(file.path(root, "Code/ETH_fig.R"))
source(file.path(root, "Code/ETH_tab.R"))
source(file.path(root, "Code/ETH_map.R"))

```

# Abstract

Maize is an important staple crop in Ethiopia with wide implications for food security and policy. In this paper we adopt the framework developed by Dijk et al. (2016) to decompose the maize yield gap in Ethiopia into (1) the technical efficiency yield gap, (2) the allocative yield gap, (3) the economic yield gap and (4) the technology yield gap. Our dataset is constructed using plot level data covering <number> Ethiopian maize plots derived from a survey of Ethiopian households. This is combined with maize specific growing information derived from crop simulation models. The results of the decomposition and econometric analysis suggest that the technology yield gap makes up the largest share of the total yield gap followed by the feasible yield gap, the technical efficiency yield gap and the economic yield gap. The results from this paper can be used to design targeted policy and and farming recommendations at the plot, farm household, local and national level.

# Introduction

Ethiopia has great agricultural potential because of its vast areas of fertile land, diverse climate, generally adequate rainfall, and large labour pool. Consequently, agriculture has remained as the foundation of Ethiopia’s economy, accounting for around 40 percent of the national income and around 80 percent of the total employment (EEA, 2015). It also supplies food to the urban areas and raw materials to the manufacturing sector. Many other economic activities depend on agriculture, including marketing, processing, and export of agricultural products. Among the major drivers of GDP growth crop production had been the second most important contributor next to construction . The agricultural sector has, therefore, been crucial for the overall performance of the Ethiopian economy and will likely continue to be the backbone of the Ethiopian economy for years to come. The sector is expected to play a central role in decreasing poverty and increasing growth in Ethiopia.

Agricultural production in Ethiopia is overwhelmingly of a subsistence nature where millions of smallholder farmers cultivate on average less than one hectare of land. About 80 per cent of smallholder production is destined for home consumption and use as seed for subsequent harvests. The smallholder sector is the main source of employment, agricultural GDP, export earnings, and food and raw materials supplies to urban areas and food industries. Crop production is by far the most dominant sub-sector, accounting for over 97 per cent of agricultural output and more than 60% of agricultural GDP (RATES, 2003). It is estimated that crop production and livestock husbandry account for over 86 per cent of the agricultural GDP (WB, 2013).

Despite this potential, agricultural productivity in Ethiopia is very low compared to other Sub-Saharan African countries. Ethiopian agriculture has remained underdeveloped due to periodic drought, soil degradation, deforestation, high population density, poor infrastructure (making it difficult and expensive to get goods to market) and low level of technology use among others. 

Maize is one of the most important staple crops in terms of calorie intake for millions of Ethiopians and is grown in many parts of the country. Because of its large geographical coverage and the scale of production, it plays a critical role in the food security of the country and particularly the poor. According to the CSA 2004/5 national survey of consumption expenditure, maize accounted for about 17 percent of the national calorie intake followed by sorghum (14 percent) and wheat (13 percent) among the major cereals (Taffesse et al., 2012). Maize is second to teff in terms of area, but it ranks first among cereals in production. According to CSA (2010) maize accounted for around 30 percent of all cereals produced in the country and more than 20 percent of the land cultivated by all cereal crops. According to the Agricultural Growth Program (AGP) Ethiopia is the fifth largest producer of maize in Africa in terms of the area devoted to maize but is second, only to South Africa, in yield and third, after South Africa and Nigeria, in production and smallholder farmers and around 9 million smallholder farmers account for more than 95 % of the total maize area and production in the country (http://ethioagp.org/maize-2/).  

Maize is grown in the mid highland areas of the country and there are huge tracts of land in all regions suitable for maize farming. Maize is grown primarily in the Amhara, Oromia and SNNPR regions of the country from which 5 zones namely West Wellega, Jimma, East Shewa, East Wellega, and West Gojjam account for over 44 percent of total smallholder production. Marketable surplus is mainly produced in these high potential zones, and the production in the other zones is mainly for subsistence. Maize production has increased in Ethiopia in recent years, due in large part to an improvement in the supply of seeds to farmers. While there have been significant gains made in maize production over the past decade, there is still a significant opportunity to further increase productivity. Ethiopia can be considered a success story for maize production as, apart from South Africa, it is the only country in Sub-Saharan Africa that shows substantial progress in maize productivity and input use (Abate et al., 2015) After a period of limited growth, yield more than doubled from around 1.5 ton/ha in 2000 to over 3 ton/ha in 2013 (FAOSTAT). The government has worked hard to ensure timely delivery of seeds and fertilizer, as well as to improve agricultural extension services and trainings for farmers. Maize yield levels are also slightly above the regional average. Despite the recent progress in productivity, yield levels in Ethiopia are still very low relative to what could be produced. According to the Central Statistical Agency of Ethiopia (CSA, 2010), the low level of crop productivity in the country is attributed to inefficient farming methods that utilize outdated crop production and protection technologies, and fragmented pieces of land holdings. Estimates indicate that the current maize yield could be doubled if farmers adopt higher quality inputs and proven agronomy best practices. Data from the Global Yield Gap Atlas (GYGA) shows that the average yield gap is on average 82% in Ethiopia. This is in strong contrast with for example Latin American countries which are able to reach around 45% of potential yield.  

Yield gap measurements are a very powerful metric to determine the absolute potential to increase production and are therefore frequently used in policy studies (Sumberg, 2012). This makes them particularly useful as a basis for forward looking scenario studies to assess future agricultural production and food security. By ‘controlling’ for the biophysical production environment, yield gaps can easily be compared across regions and countries and, therefore, provide a good indicator of relative (potential) performance. 
A limitation is that yield gap measurements offer only limited guidance in explaining the observed difference between actual and potential yield. It has been argued that it simply not cost-efficient for farmers to use the amount of inputs that are needed to reach potential. This is underscored by Lobell et al. (Lobell et al., 2009), who compared a large number of yield gap studies and find that (best-practice) farmers reach only around 80% of potential yield. The ‘exploitable’ yield gap is therefore lower than the total yield gap and is governed by economic constraints such as input prices and market access. Other important causes for the existence of yield gaps are differences in farm-level characteristics (e.g. size, education and off farm activities) and the socio-economic context (e.g. subsidies, rural infrastructure and access to finance).

The aim of this paper is the to provide a detailed analysis of the maize yield gap in Ethiopia. We decompose the conventional yield gap into four elements that capture, technical efficiency, economic efficiency, availability of technologies and an ‘unexploitable’ yield gap element.

The structure of this paper is as follows. Section 2 describes the econometric procedure used to decompose the yield gaps. Section 3 describes the data used for this study. Section 4 presents the results of our estimations and yield gap calculations. Section 5 discusses the results. Section 7 concludes.



# Policies to close the yield gap
The yield gap is defined as the difference between potential yield and actual yield. Potential yield is the maximum yield that can be produced on a piece of land given agro-climatic conditions. Potential yield can be estimated using crop models or farmer competitions. @VanDijk2017a demonstrated that the conventional yield gap (Yg) is caused by the combination of four factors, which are measured by four individual gaps: (1) technical inefficiency (technical efficiency yield gap or TEYg), (2) suboptimal allocation of resources (allocative yield gap or AYg), (3) economic constraints (economic yield gap or EYg) and (4) lack of advanced technologies (technology yield gap or TYg). This can be expressed as:

$$ Yg = TEYg + AYg + EYg + TYg $$

Dividing both sides by Yg gives the contribution of each part to the total yield gap.  

Several (sets of) policies options have been proposed to increase the productivity of smallholders. Probably, the most comprehensive discussion is provided in the 2008 World Development report [@WorldBank2008], which exclusively focussed on smallholder agriculture in the developing world. Policy solutions that are frequently proposed to improve the performance of small-scale farmers include providing credit and insurance, investing in agricultural R&D, improving extension services and providing input subsidies. In the remainder of this section, we link these and other smallholder policy options that have been proposed in the literature to the four yield gaps.    

The _technical efficiency yield gap_ is the difference between actual yield and the best-practice yield that can be achieved using the same amount of inputs and technology under similar agro-ecological conditions. The main causes for this gap are knowledge, skills and information gaps concerning optimal farm management, such as the use, combination and timing of modern inputs, which prevent farmers from reaching best-practice. The relevance of this gap is supported by a review on production constraints in Asian and African systems, which found that poor crop management resulted in 11-28% of yield losses in sub-Saharan Africa [@Waddington2010]. Similarly, a meta-analysis by @Ogundari2014 reported an average technical efficiency of 68% in Africa. Both studies indicate there is ample room to increase yield by closing the technical efficiency yield gap. 

Extension services are the main policy instrument to close the technical efficiency yield gap. The core objective of providing extension services is to address the knowledge and information gap of farmers by offering technical education and sharing information on new technologies and use of inputs [@Evenson2001]. There are many different models to provide extension services, including initiatives that simulate farmer-to-farmer knowledge exchange, such as farmers' field schools and initiatives that exchange knowledge and practices through the value chain. Investment in farmer education is another strategy that will contribute to closing the technical efficiency gap.

The _allocative yield gap_ reflects the sub-optimal allocation of resources. The existence of allocative yield gaps suggest that markets are functioning poorly or are missing completely because for some reasons farmers are not using potentially profitable inputs.[^1]  Factors that explain this behaviour operate at the demand and supply side and can be categorized as knowledge constraints, financial constraints, risk issues and information asymmetries [@Kelly2003; @Poulton2006]. A variety of policies have been proposed to ease these constraints. 

First, apart from contributing to more efficient use of inputs, extension services will also stimulate the adoption of advanced inputs by raising farmer awareness. Extensions services, if implemented effectively, will have a dual impact, contribution both to closing the technical efficiency and allocative yield gap. 

Second, the promotion of rural agro-dealer networks will simulate the use of modern inputs by improving the technical knowledge and managerial skills of traders and potentially offering credit and guarantees that solve capital constraints [@Kelly2003]. Agro-dealers, with support of seed and fertilizer suppliers, can also be instrumental in reducing the package size of inputs, which will stimulate experimentation and diversification with limited risk. 

Third, limited financial services to deal with credit and risk constraints  are frequently mentioned as a key source of market failure in African factor markets, resulting in low input use and output [@Poulton1998; @Karlan2014]. In addition to the role of agro-dealers, policy solutions that have been offered to deal credit and risk problems in rural areas are the support of micro-credit arrangements, weather insurance and mobile banking. 

Fourth, support for market information systems that strengthen the public dissemination of input and output prices will help farmers and traders to make better input use and supply decisions. 

Fifth, establishing land tenure systems will contribute to closing the allocative yield gap. Although the empirical evidence is not conclusive [e.g. @Place1993; @Place2002], some studies find a positive relationship between land rights and productivity [@Goldstein2008]. With land tenure, farmers do not have to worry about expropriation of their land, as such having more incentives to invest in land fertility and apply yield-increasing inputs. Land rights might also act as collateral making it easier to obtain credit. 

Finally, 'smart' input subsidy policies, have been proposed as a solution to overcome externalities associated with learning and other risk issues related to input use [@Morris2007]. In recent years a number of sub-Saharan African countries have introduced large-scale seed and fertilizer programs as part of agricultural development strategies to increase productivity and improve food security [see @Jayne2013 for an overview]. Malawi is probably the best-known and most studied example [e.g. @Lunduka2013a]. There is an ongoing debate about the impact and effectiveness of these input subsidy programs [@Ricker-Gilbert2013a]. In a study for three African countries, @Jayne2013a find that subsidies have a much lower impact on fertilizer use than expected because of crowding out and diversion effects. If this is properly taken into account, the cost-benefit ratio is no longer favourable. These findings, however, have been subject of debate and other studies arrive at the conclusions that, in particularly in the case of Malawi, the input subsidy program can be considered a success  [@Dorward2015; also see @Jayne2015]. 

[^1]: There is a large body that analyses the causes of missing and poorly functioning (agricultural) markets in  developing countries, which we will not repeat here [@Stiglitz1989; @Dillon2014].

The _economic yield gap_ captures the economic constraints that prevent farmers from using (more) inputs.[^2] This gap can never be completely closed because in a market economy (and in the absence of subsidies) resources will always have a price greater than zero. Several studies have point out that input costs in sub-Saharan Africa are relatively high because of a combination of transport and transaction costs in the input supply chain. @Jayne2003 provides a detailed study on the marketing costs of fertilizer in Ethiopia, Kenya and Zambia, and shows that costs can be reduced by 11-18% by means of a combination of investments in port, rail and road infrastructure, reducing port fees and high fuel taxes and  addressing uncertainties in government input distribution programs. Similarly @Minten2013  found that high transportation and transaction costs play an important role in explaining the limited use of improved seeds and chemical fertilizer in Ethiopia. The cost experienced by farmers in the “last mile(s)” between the input distribution centres and farms could raise farm gate prices by 20 to 50% in comparison to those at the distribution centre. Finally, @Antle1983 provides macro-level evidence that infrastructure development is related to agricultural productivity. 

Not surprisingly, investment in rural roads has been mentioned as a crucial policy in many studies to improve the productivity of smallholder farmers in Africa [@Jayne2010]. Reducing transaction costs requires broader policies and interventions that address macro-level and institutional problems such as reduction of distortive taxes and subsidies, streamlining regulations and trade barriers and better governance [@Jayne2010]. 

[^2]: Some researchers consider the lack of market transactions caused by high input costs, which are in turn caused by high transportation costs and thin markets, also as a market failure [see the discussion in @Jayne2003]. In our view, which is close to standard economic theory, this is not a market failure but simply efficient market behaviour because  willingness to pay is lower than the cost of supplying the input. We therefore consider high input costs and its determinants as a cause of the economic yield gap, instead of the allocative yield gap. 

The _technology yield gap_ measures the gap between potential yield and the maximum yield that is feasible with local knowledge and technologies. The main cause for this gap is (the lack of) access to and availability of appropriate technologies for smallholders in Sub-Saharan Africa [@Tittonell2013]. Examples of 11 technologies (e.g. precision agriculture, improved varieties and drip irrigation) that have the potential to increase crop yield of smallholder farmers substantially are reviewed by @Rosegrant2014. 

Closing the technology yield gap demands investment in applied research and development that facilitate the diffusion, adoption and assimilation of advanced technologies and the development of new improved varieties [@Cohen1990; @Evenson2001; @Pardey2006]. This is underlined by the high internal rate of returns that often is found in agricultural R&D impact assessments. A recent meta-analysis found a median internal rate of return of 35% for agricultural R&D in sub-Saharan Africa [@Pardey2016].  

Table 1 provides the definitions and summarises potential causes for the the various yield gaps as well as policies options that contribute to closing them. Although we linked the policies only to one yield gap, in practice, if implemented well, many of the proposed policies will, through second-order effects, contribute to closing other gaps as well. For example, by sharing information on new technologies and inputs, extension services will contribute to closing the technical efficiency yield gap. At same time, the increased skills and knowledge of farmers makes it profitable to increase the use of fertilizer even if relative prices do not change, resulting in a lower economic yield gap. Another example is the investment in rural roads, which is expected to decrease the price of fertilizer, hence contributing to closing the economic yield gap. At the same time better infrastructure will also increase access to financial services and market information, resulting in a smaller allocative yield gap. Furthermore, improved roads, will also support the diffusion of technologies and farmer-to-farmer knowledge, contributing to closing the technical efficiency gap. Finally, the investment in agricultural R&D and the adoption of advanced technology will not only help closing the technology gap but also make it more profitable to use improved seeds and fertilizer, contributing to closing the economic yield gap.

**Table 1: Yield gaps**

```{r Table-1}
p_load(readxl)
Tab_YG <- read_excel(file.path(root, "FigTabMap/Table-1.xlsx"), sheet = 1)
kable(Tab_YG)
```


# Data

The main data source is the Living Standards Measurement Study - Integrated Surveys on Agriculture (LSMS-ISA) Ethiopian Socioeconomic survey (ESS). This survey was implemented by the Central Statistical Agency of Ethiopia (CSA) and funded by the World Bank through a grant from the Bill and Melinda Gates foundation. Although part of a panel, the first wave of the ESS did not record crop harvested quantities for a majority of plots making it unsuitable for this analysis. However, the second wave (2013-2014) of the ESS provides detailed information on crop production, input use, environmental and socioeconomic variables. 

In both waves of the data respondents were visited twice as part of a post planting and post harvesting questionnaire. Through these questionnaires, detailed information at the plot level was gathered including the size of the harvest for each crop. The vast majority of plots were GPS measured, however, in some cases this was not the case and area measurements were imputed using multiple imputation and following LSMS-ISA World Bank (2014). Respondents were asked to record key production inputs including seed application, fertilizer and labour rates. Moreover, important variables regarding environmental variables, crop management practices, and the socioeconomic status of the household were recorded at the plot, household and community levels.

The longitude and latitude of each household was recorded and households were matched as part of the LSMS-ISA with climate, soil and geographical from sources including IFPRI, Worlclim, NASA and the Ethiopian Roads Agency among others. Using the longitude and latitude of the households we were able to further augment the dataset with more granular informationfrom the Africa Soil Information Service (AfSIS, http://africasoils.net) and the Global Yield Gap Atlas (GYGA, www.yieldgap.org). AfSIS provides soil quality maps for Africa at 250m spatial resolution and various depths based on 28 thousand sampling locations (Hengl et al., 2015). We use AfSIS data to derive the soil organic carbon stock and pH for the top 200cm soil layer. We also derive the water-limited potential yield from GYGA which acts as a cap on the highest yield levels that can be attained in Ethiopia from a biophysical standpoint.

```{r}
Map_GYGA_LSMS
Map_ZONE_LSMS
```



Table 1 provides summary statistics for the key variables in our study.

__Table 1: Descriptive Statistics__

```{r}
pander(sum_tab)
```

## Estimation procedures

The analytical framework in this paper follows Dijk et al. (2016), and decomposes the yield gap into its constituent parts. Central to this framework is the estimation of the frontier yield response curve. There are competing methods for estimating this function including data envelopments analysis, corrected ordinary least squares and stochastic frontiers maximum likelihood. We follow the SF ML method [@Aigner1977; @Meeusen1977] which involves specification of the production technology and a composite error term that reflects both statistical error in the model and an asymmetric inefficiency term.

\begin{equation}\label{folstandardform}
y_i = f(x_i, \beta) + \epsilon_i
\end{equation}
\begin{equation}\label{folstandardform}
\epsilon_i = v_i - u_i
\end{equation}
\begin{equation}\label{folstandardform}
v_i \sim \mathbb{N}(0, \sigma_v^2)
\end{equation}
\begin{equation}\label{folstandardform}
u_i \sim \mathbb{N}^+(0, \sigma_u^2)
\end{equation}

Where $y_i$ is the log of maize yield and $x_i$ is the log of the inputs including the nitrogen, labour and seed rates. The composite error term $\epsilon_i$ includes a truncated normal inefficiency term $u_i$ and a statistical error term $v_i$ where $u$ and $v$ are independent. With these specifications [@Aigner1977] show that the composite error term $\epsilon_i$ follows the following density

\begin{equation}\label{folstandardform}
f_\epsilon(\epsilon_i) = \frac{2}{\sigma}\phi(\frac{\epsilon_i}{\sigma})\Phi(-\frac{\lambda\epsilon_i}{\sigma})
\end{equation}

Given a suitable form for the production technology, the parameters of the stochastic frontiers model can be estimated using maximum likelihood (ML) estimation. Common production functions used in crop yield models include the Cobb Douglas and translog production functions. The translog is the more flexible of the two and nests the Cobb Douglass form allowing model selection based on a likelihood ratio (LR) test. Viewed as an extension of the cobb douglass or as a second order expansion of an unknown production function we can write

\begin{equation}\label{folstandardform}
Y_i = F(X_i, \beta) = exp \Big(\alpha + \sum_{k=1}^K \beta_k ln X_{ik} + \sum_{k=1}^K \sum_{j=1}^K \gamma_{jk} ln X_{ik} ln X_{ij} \Big)
\end{equation}

Where $Y_i$ is the output and the $X_{ik}$ are the production inputs. Taking the log of expression (6) yields:

\begin{equation}\label{folstandardform}
y_i = f(x_i, \beta) = \alpha + \sum_{k=1}^K \beta_k x_{ik} + \sum_{k=1}^K \sum_{j=1}^K \gamma_{jk} x_{ik} x_{ij} + \epsilon_i
\end{equation}

Where the log of the output is written as $y_i$ and the log of the inputs are written as $x_i$. Expression (7) has the familiar translog form, or, setting setting $\gamma_{jk} = 0$, the Cobb Douglas form. Expression (7) is then augmented with a composite error term $\epsilon_i$ comprised of the truncated normal technical inefficiency and the symmetric error term $\epsilon_i = v_i - u_i$. In this form it is possible to carry out SF ML estimation as described in (1) - (5) above. We can also include environmental variables such as the slope and elevation of the plot yielding the following model in the translog case:

\begin{equation}\label{folstandardform}
y_i = \alpha + \sum_{k=1}^K \beta_k x_{ik} + \sum_{k=1}^K \sum_{j=1}^K \gamma_{jk}x_{ij} x_{ik} + W_i \theta + \epsilon_i
\end{equation}

Where $W_i$ are the environmental variables and $\theta$ is a vector of parameters.

As our interest is also in the determinants of the inefficiency term we allow the mean of the pretruncated normal inefficiency term to depend on exogenous determinants of inefficiency such as the education, age or gender of the farmer or access to markets and extension services. In this case equation 4 becomes:

\begin{equation}\label{folstandardform}
u_i \sim \mathbb{N}^+(z_i\delta, \sigma_u^2)
\end{equation}

Where $z_i$ is a vector of exogenous determinants of inefficiency. Estimation is carried out by maximum likelihood using the density for the error term $\epsilon_i$ and incorporating this new term. The resulting coefficients $\delta$ are not directly interpretable but the marginal effects of the exogenous determinants of technical inefficiency on the mean of technical inefficiency $u_i$ can be derived. We follow Kumbhakar and Sun (2013) and calculate the marginal effects on the conditional mean of technical inefficiency in keeping with the Jondrow et al (1982) estimator.

Finally, our framework for decomposing maize yields includes a measure of the economically optimal level of maize yield, The yield at which profits are maximized. This level can be identified as the point at which the relative market price of the input to the output is tangent to the frontier yield response function. The tangent or marginal physical product (MPP) of (7) with respect to one of the inputs $k$ is

\begin{equation}\label{folstandardform}
MPP = \frac{Y}{N}\Big( \beta_k + 2\beta_k + \sum_j ln x_j \Big)
\end{equation}

which we equate with the relative price $\frac{Pn}{Pm}$ and solve for $N*$

\begin{equation}\label{folstandardform}
\frac{Y}{X_k}\Big( \beta_0 + 2\beta_k x_k + \sum_j ln x_j \Big) - \frac{Pn}{Pm} = 0
\end{equation}

The value of nitrogen that solves this relationship is the economically optimal nitrogen input level. However, this value cannot be found analytically. Instead, it is necessary to use numerical methods to solve for the optimal use of inputs.

## Endogeneity 

Still lots to add

### testing for endogeneity

easy using control function approach

- We note from Angrist and Krueger that an OLS (instead of tobit) first stage is OK, and Wooldridge also makes this point, although neither seem to explain why. Should also mention tobit + literature on average production functions arguing for tobit

### Accounting for endogeneity

Harder using LIML approach, or two step LIML or corrected 2sls

- in progress
- mainly based on Amsler et al. (2016)
- seems to work actually quite well.

## Results

We initially test the suitability of a 
The resulting SF ML estimates are presented in table 1 for four models. Model 1 shows the coefficient estimates for the core translog production function terms. Model 2 extend this to include the exogenous determinants of inefficiencies. Models 3 and 4 do likewise and also include the environmental variables.

__Table 2: Translog production function__
```{r}
pander(results_tab)
```

The raw coefficients of the exogenous determinants of technical inefficiency cannot be interpreted as elasticities. However, the marginal effects of each variable can be computed following Kumbhakar & Sun (2013). These are observation specific and a concise estimate of the marginal effects is presented in table 2 in the form of the average partial effects (APE) for models 3 and 4 respectively. 

__Table 3: Marginal effects (APE) of exogenous determinants of inefficiency__
```{r}
pander(ME_tab)
```


__Figure 3: Potential yield__

```{r Fig_2}
Fig_waterfall
```


__Table 4: __

```{r}
pander(ZonalYieldGap_l_sh)
```


