---
title: "Preliminary Yield Gap Analysis ETH"
author: "Michiel van Dijk & Tom Morley"
date: "14 December 2015"
output: beamer_presentation

---
```{r, echo=FALSE}
# Get workspace with all info
dataPath <- "D:\\Data\\IPOP\\SurveyData\\"
wdPath <- "D:\\Dropbox\\Michiel_research\\2285000066 Africa Maize Yield Gap"
setwd(wdPath)
load(".\\Analysis\\ETH\\Report\\WS.RData")
```

# Contents


- Data description
- Results
- Work in progress

# Data

* LSMS-ISA for ETH, 2013 only
    + Inputs: Capital, labour, nitrogen, control variables, etc
    + Outputs: Maize yield
    + Prices: Maize and fertilizer
* GYGA results for TZA
* AFSIS: pH and SOC

# LSMS-ISA

```{r, echo=FALSE}
p4
```

# GYGA

```{r, echo=FALSE}
p3
```

# Model
\tiny
```{r table2, results='asis', message=FALSE, echo=FALSE} 
library(xtable) 
res.table <- xtable(sfatable, caption = 'stochastic frontier', table.placement ="")
print(res.table, caption.placement = "top", comment = FALSE) # scalebox=0.7 can be used to shrink text as alernative for "tiny".


# Alternative using Pandoc - to explore
#```{r, kable, echo=FALSE} 
# http://stackoverflow.com/questions/16424545/reduce-cell-width-and-font-size-of-table-using-pandoc-table
#options(digits=2)
#library(knitr)
#kable(sfatable)
#```
``` 

# Distribution of yield gaps
```{r, echo=FALSE}
p2
```

# Potential to increase maize production in ETH
```{r, echo=FALSE}
gg
```

# Potential to increase maize production in ETH
```{r, echo=FALSE}
gg
```

# Work in progress
* Exploratory analysis of data 
* Extend production function (e.g. soil variables)
* Add explanatory (z) variables
* Regress EYG, TYG and YG on explanatory factors
* Submit abstract to AAAE (23-26 September, Addis Ababa)
