---
title: "ETH Yield Gap Analysis"
author: "Michiel van Dijk & Tom Morley"
date: "27 February 2017"
output: beamer_presentation
#output:
#  revealjs::revealjs_presentation:
#    incremental: true
#    theme: moon
#    highlight: zenburn
#    center: true
#    transition: convex
#    reveal_options:
#      slideNumber: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# set project root
library(rprojroot)
library(pander)
root <- find_root(is_rstudio_project)
# source(file.path(root, "Code/ETH_tab.R"))
source(file.path(root, "Code/ETH_fig.R"))
```

## What's new since last time

- A more flexible translog function allowing for interactions between inputs
- Control function approach to incorporate a feedback loop between nitrogen and yield
- numerical methods to find optimum nitrogen use
- GYGA variables 

## Yield gap framework

@michiel insert picture

## Data

- Two (now three) LSMS-ISA surveys 2011 and 2013 (and as of last week 2015)
- But due to a mistake in recording production in 2011 we only analyse 2013
- Yield defined as production/plot size => different from FAOSTAT/GYGA 

## Summary statistics

\small
```{r,  results="asis"}
# http://tex.stackexchange.com/questions/10863/is-there-a-way-to-slightly-shrink-a-table-including-font-size-to-fit-within-th
suppressMessages(library(stargazer))
dbsum <- readRDS(file.path(root, "Cache/dbsum.rds"))
stargazer(as.data.frame(dbsum),  type = "latex", digits=2, header=FALSE, title = "Summary statistics")
```

## Methodology

- Two stage estimation approach
+ First stage Tobit model to avoid feedback loop/confounding 
+ Second stage stochastic frontiers estimation (bootstrapping SEs -> nonlinear function)
+ Third stage policy variables

\normalsize


## LSMS-ISA and GYGA

@ Michiel add in the GYGA map
```{r GYGA_map}
#GYGA_LSMS
```


## Data issues

- How to solve missing yield potential data?
    + Average values are too low, maximum is too high.
- To calculate feasible yield gap information is needed on the use of inputs when costs are zero:
    + Level of nitrogen when yield diminishes (per region)
    + Use of labour and capital at this point
- Information on nitrogen/fertilizer use can be taken from experimental plot data => literature survey.
- Use of labour and capital demands assumptions (e.g. 10% increase)


## Stochastic frontier analysis

\tiny
```{r sfa}
results <- readRDS(file.path(root, "Cache/results_table.rds"))
row.names(results) <- NULL
pander(results)
#, caption="Stochastic Frontier results") 
```
\normalsize


## Yield gap estimations
\small 
```{r relative_YG}
knitr::kable(ZonalYieldGap_l, caption = "Relative yield gap", digits = 0)
```
\normalsize


\normalsize


## Closing the yield gap

```{r Gap}
Fig_waterfall
```


## Policies to close the yield gap

@ Michiel policy framework picture


## Explaining the TEYG
\tiny
```{r z_table}
# knitr::kable(ztable, caption = "Second stage analysis", digits = 2)
```
\normalsize



## Next steps

- Calculate elasticities of second stage variables to get a better interpretation
- Think about how to link policies to yield gaps
- finish writing paper
- Third wave of data???? but ..... time constraints

