---
title: "Ethiopia Yield Gap Analysis"
author: "Michiel van Dijk & Tom Morley"
date: "10 April 2017"
output: beamer_presentation
#output:
#  revealjs::revealjs_presentation:
#    incremental: true
#    theme: moon
#    highlight: zenburn
#    center: true
#    transition: convex
#    reveal_options:
#      slideNumber: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# set project root
library(rprojroot)
library(pander)
root <- find_root(is_rstudio_project)
# source(file.path(root, "Code/ETH_tab.R"))
source(file.path(root, "Code/ETH_fig.R"))
# source(file.path(root, "Code/ETH_map.R"))
```

## Outline of analysis

- Decomposition of maize yield gap in Ethiopia
- A flexible translog production allowing for interactions between inputs
- Numerical methods for finding optimal nitrogen use 
- GYGA environmental control variables 
- Heterogeneity not dealt with due to data limitations
- Endogeneity

## Yield gap framework

## LSMS-ISA and GYGA

## Data

- Two (now three) LSMS-ISA surveys 2011 and 2013 (and as of a month ago, 2015 - 2016)
- But due to a mistake in recording production in 2011 we only analyse 2013
- Yield defined as production/plot size => different from FAOSTAT/GYGA 

## Summary statistics

\small
```{r,  results="asis"}
# http://tex.stackexchange.com/questions/10863/is-there-a-way-to-slightly-shrink-a-table-including-font-size-to-fit-within-th
suppressMessages(library(stargazer))
dbsum <- readRDS(file.path(root, "Cache/dbsum.rds"))
stargazer(as.data.frame(dbsum),  type = "latex", digits=2, header=FALSE, title = "Summary statistics")
```
\normalsize


## Methodology

- Stochastic frontiers analysis
- Includes an asymmetric error component which accounts for plot level inefficiency.
- Estimated by maximum likelihood
- Altered density of error term to allow exogenous determinants of inefficiency such as extension services.
- Marginal effects of exogenous determinants of inefficiency


## Stochastic frontier analysis

\tiny
```{r ech=FALSE, message=FALSE, warning=FALSE, sfa}
source(file.path(root, "Code/production_function_tests.R"))
pander(results_tab)
```
\normalsize


## Yield gap estimations

\small 
```{r relative_YG}
ZonalYieldGap_l_sh <- ZonalYieldGap_l_sh %>%
  rename(FYG = EUYG)
ZonalYieldGap_l_sh$Zone <- gsub("BENSHANGULGUMUZ", "BG", ZonalYieldGap_l_sh$Zone)
names(ZonalYieldGap_l_sh) <- c("ZONE" ,"TEYG", "AYG", "EYG", "TYG", "YG")
knitr::kable(ZonalYieldGap_l_sh, caption = "Relative yield gap", digits = 0)
```
\normalsize

## Closing the yield gap

```{r Gap}
Fig_waterfall
```

## Explaining the TEYG

- marginal effects of exogenous determinants of inefficiency.

\small
```{r z_table}
pander(ME_tab)
```
\normalsize

## Remaining issues

- Take cap values for nitrogen from nutrient gap project? Also useful for other countries. 

- Reasonable increases in other variables such as labour and seed rates and dummy variables. Plus anything else we might expect?

- Missing yield potential data - currently taking the national maximum, average values are too low, maximum is too high!

## Current assumptions

- Nitrogen at 400 kg/ha - link to nutrient gap project
- Seed rate increased by 50%
- labour increased by 50%
- dumoxen = 1, all plots use animal traction.
- Other suggestions? Other dummies are Ph, crop count.
- National maximum for water limited potential yield = 18072

## Endogeneity

- Idea of a feedback loop between the level of nitrogen applied and the yield.

- Agronomic interpretation: Higher yield in good soils, but Nitrogen also works better in good soil resulting in confounding

- Policy interpretation: Farmers listen to extension agents and see improvements spurring on more improvements.

- Bottom line: failing to account for endogeneity in survey data results in biased estimates.

- For nitrogen this is most likely biased up implying that we overstate the impact of nitrogen on yields. And there is a large literature identifying the endogeneity of nitrogen so we should include this.

- Solution in stochastic frontiers analysis is to use a 2 stage limited information ML technique.

## first stage

\tiny
```{r echo=FALSE}
pander(readRDS(file.path(root, "Cache/endog1.rds")))
```
\normalsize

- R-squared on first stage is only 0.238, but this is OK because LIML performs well in the presence of weak instruments. Including other variables that affect the demand for nitrogen will also increase this.

## Second stage

\tiny
```{r echo=FALSE}
pander(readRDS(file.path(root, "Cache/endog2.rds")))
```
\normalsize

- Bootstrapped second stage SEs because v term is an estimate of the true error.

## SF ML vs LIML with endogeneity

\tiny
```{r echo=FALSE}
pander(readRDS(file.path(root, "Cache/endog3.rds")))
```
\normalsize

## Wrap up

- New and better translog function means more flexibility of inputs
- Optimal nitrogen levels calculated per farmer using their actual values of other inputs
- Marginal effects of exogenous variables calculated giving a policy interpretation
- Identified bias introduced due to endogeneity but this needs to be extended somewhat.
